---
title: "BankDataMain"
format: html
editor: visual
---

```{r echo=FALSE, message=FALSE, warning= FALSE}

library(tidyverse)
library(tidymodels)
library(lubridate)
library(scales)
library(randomForest)
library(caret)
library(class)
library(e1071)
library(ggcorrplot)
library(rstatix)
```


```{r message=FALSE , echo = FALSE}
Data <- readRDS("cleaned_data.rds")
Dataframe_base <- read_csv2("bank-full.csv")
```

```{r}
colSums(is.na(Data)) #Sprawdzenie braków danych
```

## Distribution of customers' ages: What is the distribution of customers' ages? Are there any dominant age groups?

```{r}
Data %>%
  ggplot(aes(x = age, fill = age)) +
  geom_bar() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Number of clients") +
  theme(legend.position = "none")
```

Conclusions: The data has lots of "Middle-Aged Adults" and "Older Adults," showing they're focused on. But there aren't as many "Younger Adults" or "Seniors," so they might not have been as important in the campaign.

## Relationship between age and deposit subscription: Is there a connection between age and subscribing to a deposit? Are certain age groups more inclined towards subscribing?

```{r}
Data %>%
  group_by(age) %>%
  ggplot(aes(x = age, fill = y)) +
  geom_bar(position = "fill") +
  labs(x = NULL, y = "Proportion of term deposit subscriptions", fill = "Has the client subscribed a term deposit?") +
  theme(legend.position = "bottom")
```

Conclusions: There appears to be a correlation between age and subscribing to a deposit. Younger Adults and Seniors show a higher inclination towards subscribing to deposits compared to Middle-Aged and Older Adults.

## Annual balance and subscription: Is the average annual balance of clients connected to deposit subscription?

```{r}
Data %>%
  group_by(y) %>%
  summarise(avg_balance = mean(balance)) %>%
  ggplot(aes(x = y, y = avg_balance, fill = y)) +
  geom_bar(stat = "identity")
```

## Contact duration and subscription: Does the duration of the last contact relate to deposit subscription?

```{r}
Data %>%
  mutate(
    duration = duration / 60,
    duration_group = cut(duration, breaks = c(-1, 5, 15, max(duration) + 1), labels = c("Short", "Long", "Very long")), right = TRUE
  ) %>%
  group_by(duration_group) %>%
  ggplot(aes(x = duration_group, fill = y)) +
  geom_bar(position = "dodge")

```

Conclusions: There seems to be a relationship between the duration of the last contact and deposit subscription. Longer contacts tend to have a higher number of deposit subscriptions compared to shorter contacts.

## Is there a correlation between outcome of the prevoius campaign and the subscription to a term deposit in the current campaign?

```{r}
Data %>%
  ggplot(aes(x = poutcome, fill = y)) +
  geom_bar(position = "dodge")

```

Conclusions: This implies that the outcome of the previous campaign might influence the subscription behavior in the current campaign, suggesting a correlation between these variables.

## Is the marital status of clients (Marital) correlated with the subscription to a term deposit (Y)?

```{r}
Data %>%
  ggplot(aes(x = marital, fill = y)) +
  geom_bar(position = "fill")

```

Conclusions: There is not a trend indicating a strong correlation between marital status and subscription to a term deposit

## How does the job status relate to the subscription of a term deposit?

```{r}
Data %>%
  ggplot(aes(x = job, fill = y)) +
  geom_bar(position = "dodge")

```

Conclusions: The data suggests that job status might have some influence on the subscription to a term deposit, as evidenced by varying subscription counts among different job categories. Yet, this relationship isn't straightforward or consistent across all job types, indicating that other factors might also play a role in determining subscription behavior.

## How does the education level relate to the subscription of a term deposit?

```{r}
Data %>%
  ggplot(aes(x = education, fill = y)) +
  geom_bar(position = "fill")
```

Conclusions: The data suggests a correlation between education level and the subscription to a term deposit. Higher education levels, specifically tertiary education, seem to be associated with a higher likelihood of subscribing to a term deposit.


```{r}
# macierz korelacji

Dataframe_numeric <- Dataframe_base[sapply(Dataframe_base, class) != "character"]

corr <- cor_mat(Dataframe_numeric)
p_mat <- cor_pmat(Dataframe_numeric)
ggcorrplot(corr, p.mat = p_mat,lab = T)
```

```{r}
plyr::count(Data, "y")

ggplot(Data, aes(x=y))+
  geom_bar(fill = "lightgreen")
```

## Modeling

```{r}

data <- readRDS("cleaned_data.rds")

set.seed(2000)
splited_data <- initial_split(data, 0.9)
train_data <- training(splited_data)
test_data <- testing(splited_data)
```

```{r}
# Random Forest
model_random_forest <- randomForest(
  formula = y ~ age + education + poutcome + duration,
  data = train_data,
  type = "classification"
)

summary(model_random_forest)

predictions <- predict(model_random_forest, test_data)
pred_df <- cbind(test_data, predictions)

pred_df$predictions <- ifelse(pred_df$predictions > 0.5, TRUE, FALSE)

conf_matrix_model_random_forest <- confusionMatrix(data = as.factor(pred_df$predictions), reference = as.factor(pred_df$y))

conf_matrix_model_random_forest
```

```{r}
# Linear SVM 
train_data_svm <- train_data %>% select(age, education, poutcome, y) %>% mutate(y = as.integer(y))
train_dummies <- dummyVars("~.",
            data = train_data_svm)
train_data_svm <- predict(train_dummies, train_data_svm)


test_data_svm <- test_data %>% select(age, education, poutcome, y) %>% mutate(y = as.integer(y))
test_dummies <- dummyVars("~.",
            data = test_data_svm)
train_data_svm <- predict(test_dummies, test_data_svm)

model_svm <- svm(y ~ ., data = test_data_svm, kernel = "linear")
predictions <- predict(model_svm, test_data_svm[, -4])
predictions <- as.integer(ifelse(predictions > 0.5, 1, 0))

result <- cbind(test_data_svm, predictions)
conf_matrix_model_svm <- confusionMatrix(data = as.factor(result$predictions),
                                         reference = as.factor(result$y))
conf_matrix_model_svm
```


